### 2017-07-09 submitted/response.txt

First, we would like to thank all reviewers for their insightful comments and
suggestions. Besides fixing typos and minor cosmetic issues, we also tried our
best to address the comments. Responses to each individual reviewers are listed
below.

To reviewer 1:

With respect to the edit distance, that's precisely the idea behind augmented
transitions. Type-1 transitions *inserts* packets to the sniffer trace while
Type-2 transitions *removes* packets from the sniffer trace. The number of
augmented transitions (Type-1 and Type-2) represents the "distance" between the
mutation trace and the original sniffer trace. The gist of the NumMissing
heuristic we proposed is exactly to limit the "distance" (hence edits) to the
sniffer trace.  In other words, the "distance" between the mutation trace and
sniffer trace represents the confidence of protocol correctness.

To reviewer 2:

With respect to the verification coverage future work, the current proposal
provides ways to analyze sniffer traces that are already being collected. While
this is useful in revisiting existing traces which are mostly likely only spot
checked, the verification is still limited to a static sets of sniffer trace
which may only cover the "normal" path of a protocol excitation. Consider the
case study of the game controller protocol, the rate control related violations
were only detected in a tiny fraction of the traces since the violation requires
non-trivial packet loss patterns to manifest itself. The idea behind our
verification coverage proposal is to proactively control the process of
experiment set up and sniffer trace collection to increase the chance of each
protocol state being visited, hence increasing the  probability of violation
detection.

To reviewer 3:

With respect to the fault model, in this paper we only consider full packet
loss, which reflects most practical sniffer's behavior. In most wireless
communications protocols, per-packet checksum is used and a packet that fails
the checksum will be dropped by the network driver.

With respect to the detailed NS-3 experiment setup, we added a paragraph in
Section 5.1.1 to explain the packet loss model we used in the experiments.
Basically, we added a new propagation loss component to NS-3 so that we can
control per-packet losses between each individual pair of devices.

### 2017-11-01 related.tex
working on related

\textbf{Hidden Markov Model (HMM) Approach.} When considering the whole system
under test (both DUT and endpoint), the sniffer only captures a subset of the
all the packets (events).  This is similar to the event sampling problem in
runtime
verification~\cite{bonakdarpour2011sampling,hauswirth2004low,arnold2008qvm,fei2006artemis,basin2012monitoring}.
Stoller \textit{et al}~\cite{stoller2011runtime} used HMM-based state estimation
techniques to calculate the confidence that the temporal property is satisfied
in the presence of gaps in observation.

While it seems possible to adapt the
method in~\cite{stoller2011runtime} to our problem, we note several advantages
of our approach. First, the
automatically augmented state machine precisely encodes the protocol
specification and the uncertainty. This is intuitive to design and natural for
reporting the evidence for a trace being successful. We do not require a user
to specify the number of states of the underlying HMM, or accurately provide
underlying probabilities. Second, we use timed automata to monitor the timing
constraints which are common in wireless protocols. It may be non-trivial to
encode such timing information in HMM. Finally, we can exploit domain knowledge
to devise effective pruning heuristics to rule out unlikely sequences during the
exhaustive search.

>>>>

\textbf{Hidden Markov Model (HMM) Approach.} When considering the whole system
under test (both DUT and endpoint), the sniffer only captures a subset of the
all the packets (events).  This is similar to the event sampling problem in
runtime
verification~\cite{bonakdarpour2011sampling,hauswirth2004low,arnold2008qvm,fei2006artemis,basin2012monitoring,sistla2011runtime}.
In particular, Stoller \textit{et al}~\cite{stoller2011runtime} used HMM-based
state estimation techniques to calculate the confidence that the temporal
property is satisfied in the presence of gaps in observation. HMM-based
approaches proposed in~\cite{stoller2011runtime} suffer from overhead of both
running time and memory consumption. Optimizations 
such as approximate precomputed tables~\cite{bartocci2012adaptive} and particle
filtering~\cite{kalajdzic2013runtime} were later proposed.

In our problem, there is not only gaps in the observation (packets missed by the
sniffer), but also uncertainty regarding observed events, which lies in whether
a packet in the sniffer trace was received by its destination device. This
unique challenge makes it infeasible to directly apply the HMM-based approaches
proposed in~\cite{stoller2011runtime}.
Besides, there are several advantages of our proposed augmented transition
approach. First, the automatically augmented state machine precisely encodes the
protocol specification and the uncertainty. This is intuitive to design and
natural for
reporting the evidence for a trace being successful. We do not require a user
to specify the number of states of the underlying HMM, or accurately provide
underlying probabilities. Second, we use timed automata to monitor the timing
constraints which are common in wireless protocols. It may be non-trivial to
encode such timing information in HMM. Finally, we can exploit domain knowledge
to devise effective pruning heuristics to rule out unlikely sequences during the
exhaustive search.

\textbf{Edit Distance.} The closet work to ours is the Edit distance based approach
to monitor STL properties of real time systems~\cite{jakvsic2016quantitative}.
